\documentclass[11pt]{book} 

\tolerance=600

% for \begin{center}, etc.
\usepackage[margin=1.0in]{geometry}

% all kinds of math macros
\usepackage{amsmath}
\usepackage{amssymb}

% eps figures
\usepackage{epsfig}

% chapter title styles
\usepackage[Sonny]{fncychap}
\ChNameVar{\LARGE}
\ChTitleVar{\LARGE\sl}

% part page style see
% http://tex.stackexchange.com/questions/6609/problems-with-part-labels-using-titlesec
\usepackage{titlesec}

\titleformat{\part}[display]
   {\Huge\filcenter}
   {{\partname{}} \thepart}
   {0em}
   {\hrule}


% hyperlinks -- load after fncychap
\usepackage{hyperref}

% color package
\usepackage[usenames]{color}

% make the MarginPars look pretty
\setlength{\marginparwidth}{0.75in}
\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily
\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

% to increase the likelihood that floats will occur "here" when you
% want them to
\renewcommand{\floatpagefraction}{1.0}
\renewcommand{\topfraction}{1.0}
\renewcommand{\bottomfraction}{1.0}
\renewcommand{\textfraction}{0.0}

% number subsubsections and put them in the TOC
\setcounter{tocdepth}{3}
\setcounter{secnumdepth}{3}

% custom hrule for title page
\newcommand{\HRule}{\rule{\linewidth}{0.125mm}}


% short table of contents
\usepackage{shorttoc}

% spacing in the table of contents
\usepackage[titles]{tocloft}

\setlength{\cftbeforechapskip}{2ex}
\setlength{\cftbeforesecskip}{0.25ex}

% For splitting up lists into multitple columns
\usepackage{multicol}

% don't put a header on blank pages, see
% http://www.latex-community.org/forum/viewtopic.php?f=4&p=51559
% change ``plain'' to ``empty'' to eliminate the page number
\makeatletter
\renewcommand*\cleardoublepage{\clearpage\if@twoside
\ifodd\c@page\else
\hbox{}
\thispagestyle{empty}
\newpage
\if@twocolumn\hbox{}\newpage\fi\fi\fi}
\makeatother


% don't make the chapter/section headings uppercase.  See the fancyhdr
% documentation (section 9)
\usepackage{fancyhdr}
\renewcommand{\chaptermark}[1]{%
 \markboth{\chaptername
\ \thechapter.\ #1}{}}

\renewcommand{\sectionmark}[1]{\markright{\thesection---#1}}

\graphicspath{{CastroVerification/}{ConvertCheckpoint/}{Parallel/}{Scaling/}{Visualization/}}

% skip a bit of space between paragraphs, to enhance readability
\usepackage{parskip}



% special fraction
\newcommand{\sfrac}[2]{\mathchoice
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
  {#1\!/#2}}

\def\Ab {{\bf A}}
\def\eb {{\bf e}}
\def\Fb {{\bf F}}
\def\gb {{\bf g}}
\def\Hb {{\bf H}}
\def\ib {{\bf i}}
\def\Ib {{\bf I}}
\def\Kb {{\bf K}}
\def\lb {{\bf l}}
\def\Lb {{\bf L}}
\def\nb {{\bf n}}
\def\Pb {{\bf P}}
\def\Qb {{\bf Q}}
\def\rb {{\bf r}}
\def\Rb {{\bf R}}
\def\Sb {{\bf S}}
\def\ub {{\bf u}}
\def\Ub {{\bf U}}
\def\xb {{\bf x}}

\def\dt       {\Delta t}
\def\omegadot {\dot\omega}

\def\inp  {{\rm in}}
\def\outp {{\rm out}}
\def\sync {{\rm sync}}

\def\half   {\frac{1}{2}}
\def\myhalf {\sfrac{1}{2}}
\def\nph    {{n+\myhalf}}

% codes
\newcommand{\castro}{{\sf Castro}}
\newcommand{\maestro}{{\sf Maestro}}
\newcommand{\boxlib}{{\sf BoxLib}}
\newcommand{\microphysics}{{\sf Microphysics}}
\newcommand{\yt}{{\sf yt}}
\newcommand{\amrvis}{{\sf Amrvis}}
\newcommand{\visit}{{\sf VisIt}}

\usepackage{listings}

\definecolor{gray}{rgb}       {0.8,0.8,0.8}
\definecolor{light-blue}{rgb} {0.8,0.8,1.0}
\definecolor{light-green}{rgb}{0.8,1.0,0.8}
\definecolor{light-red}{rgb}  {1.0,0.9,0.9}

\lstset{
  basicstyle=\small\ttfamily,%
  frame=shadowbox,%
  rulesepcolor=\color{gray},%
  backgroundcolor=\color{white}%
}


%------------------------------------------------------------------------------
\begin{document}

\frontmatter

\begin{titlepage}
\begin{center}
\ \\[3in]
{\sf \Huge Microphysics} 

\begin{minipage}{5.5in}
\HRule\\[2mm]
\centering
{\Large \em A collection of astrophysical microphysics routines \\ with interfaces to the BoxLib codes}

\HRule
\end{minipage}

\ \\[1 in]
{\sf \huge User's Guide}

\vfill

{\large \today}
\end{center}

\end{titlepage}


\shorttoc{Chapter Listing}{0}

\setcounter{tocdepth}{2}
\tableofcontents

\clearpage

\chapter*{Preface}
\chaptermark{Preface}
\addcontentsline{toc}{chapter}{preface}

Welcome to the \microphysics\ User's Guide!

In this User's Guide we describe the microphysics modules designed to
support \boxlib\ codes such as \castro\ and \maestro. These all have a
consistent interface and are designed to provide the users of those
codes an easy experience in moving from the barebones microphysics
modules provided in those codes. For the purposes of this user's
guide, the microphysical components we currently deal with are the
equation of state (EOS) and the nuclear burning network.

\microphysics\ is not a stand-alone code. It is intended to be used in
conjuction with \boxlib\ codes and so we do not provide support for
running these codes separately.  However, in many cases we will
provide test modules that demonstrate a minimal working example for
how to run the modules.

\clearpage

\mainmatter


\chapter{The Basics}

Getting started with \microphysics is straightforward. Because the
modules here are already in a format that the \boxlib\ codes
understand, you only need to provide to the code calling these
routines their location on your system. The code will do the rest. To
do so, define the {\tt MICROPHYSICS\_DIR} environment variable, either
at a command line or (if you use the {\tt bash} terminal) through your
{\tt $\sim$/.bashrc} profile, e.g.:
\begin{equation*}
  \texttt{export MICROPHYSICS\_DIR = /path/to/Microphysics}
\end{equation*}
The calling code will know that when you provide it an {\tt EOS\_dir}
and {\tt Network\_dir}, that it should look in this repository and not
in its own microphysics location.

\chapter{Data Structures}
\label{sec:data_structures}

All of the routines in this software package are standardized so that
you interact with them using the same type of data structure. We call
this the {\tt eos\_t\_vector}. This is a Fortran 90 derived data type
(which, for our purposes, is analogous to a C++ struct). It is a
collection of data specifying the microphysical state of the fluid
that we are evaluating.  This derived type has many components, and in
the Fortran syntax we access them with the \% operator. For a
particular instantation named {\tt eos\_state}, the most important
data is the following. Most of these components are one-dimensional
(rank-1) Fortran arrays that consist of pointers to the corresponding
data, except for a few with additional information.
\begin{itemize}
  \item {\tt eos\_state \% rho}: density.
  \item {\tt eos\_state \% T}: temperature.
  \item {\tt eos\_state \% p}: pressure.
  \item {\tt eos\_state \% e}: internal energy.
  \item {\tt eos\_state \% h}: enthalpy.
  \item {\tt eos\_state \% s}: entropy.
  \item {\tt eos\_state \% xn}: mass fractions of species (this is a
    2D array, where the second index has the same number of components
    as there are fluid species)
  \item {\tt eos\_state \& aux}: any auxiliary variables carried with
    the fluid (also a 2D array, where the second index has as many
    components as there are auxiliary variables)
\end{itemize}
There is a lot more information that can be saved here, such as the
partial derivatives of the thermodynamic state variables with respect
to each other. To see a complete list, examine the {\tt eos\_type.f90}
file inside the code calling \microphysics (e.g. {\tt
  Castro/EOS/eos\_type.f90}.)

\chapter{Equations of State}

\input EOS/eos.tex


\chapter{Nuclear Networks}

\input networks/networks.tex


%------------------------------------------------------------------------------
\backmatter

\renewcommand\bibname{References}
\addcontentsline{toc}{chapter}{References}
\bibliographystyle{plain}
\bibliography{refs}

\end{document}
